Solving the power flow problem on integrated transmission-distribution networks: a review and numerical assessment

### 2020-10-28 IEEEtran/extra.tex
Draft_before_revision

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal]{article}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

%\usepackage[margin=1.5in]{geometry}
\usepackage{algorithm}
\usepackage{algpseudocode}
\makeatletter 
\renewcommand\thealgorithm{\thesection.\arabic{algorithm}} 
\@addtoreset{algorithm}{section} 
\makeatother
\usepackage{color,soul} %highlight 
% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
%\fi
% % graphicx was written by David Carlisle and Sebastian Rahtz. It is
% % required if you want graphics, photos, etc. graphicx.sty is already
% % installed on most LaTeX systems. The latest version and documentation
% % can be obtained at: 
% % http://www.ctan.org/pkg/graphicx
% % Another good source of documentation is "Using Imported Graphics in
% % LaTeX2e" by Keith Reckdahl which can be found at:
% % http://www.ctan.org/pkg/epslatex
% %
% % latex, and pdflatex in dvi mode, support graphics in encapsulated
% % postscript (.eps) format. pdflatex in pdf mode supports graphics
% % in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% % that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% % not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% % which can result in "jaggedy"/blurry rendering of lines and letters as
% % well as large increases in file sizes.
% %
% % You can find documentation about the pdfTeX application at:
% % http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
% \usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools} %includes and extends `amsmath'. Needed for dcases
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

\usepackage{blkarray, bigstrut} %
\usepackage{kbordermatrix}% http://www.hss.caltech.edu/~kcb/TeX/kbordermatrix.sty
% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithm}
\usepackage{algpseudocode}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
 \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf,position=top]{subfig}
%\else
 %\usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
% \usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e
\usepackage{epstopdf}

\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.
\usepackage[hidelinks]{hyperref} % this should allow links (to sections but also to equations, figures, references) in pdf.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% Allow page breaks inside equations (or some other math environments). The optional argument ranges from 1 to 4. The higher the number, the more permissive you are of pagebreak
%\allowdisplaybreaks[1]

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].
%\usepackage{caption}
%\usepackage{subcaption}
\usepackage{tabularx}

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

% packages needed for plotting results
\usepackage{tikz}
\usetikzlibrary{trees}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.15}
\usepackage{multirow}
\usepackage{booktabs} %for fancy tables
\usepackage{adjustbox}

\usepackage{todonotes} 
% \usepackage[disable]{todonotes} %Hides all(!) todonotes
% todonotes appearance
\newcommand{\addref}{\todo[color=red!40,caption={}]{Add reference.}}
\newcommand{\rewrite}[1]{\todo[inline,color=yellow!40,caption={}]{#1}} % the caption={} is needed if you want to include itemize and such in the todonotes, see: https://tex.stackexchange.com/questions/54064/itemize-inside-todonote
\newcommand{\question}[1]{\todo[inline,color=blue!40,caption={}]{#1}}
\newcommand{\includes}[1]{\todo[inline,color=black!10,caption={}]{#1}}
% % Don't print any [non-inline] notes (will still show up in list of todo notes, to disable this, see: http://tex.stackexchange.com/questions/287394/todonotes-show-inline-hide-all-others)
% \makeatletter
% \renewcommand{\@todonotes@drawMarginNoteWithLine}{}
% \makeatother

\makeatletter
\renewcommand*\env@matrix[1][\arraystretch]{%
  \edef\arraystretch{#1}%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{*\c@MaxMatrixCols c}}
\makeatother

% Thick vertical line (") and thick horizontal line (\thickhline) for table
\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\usepackage[per-mode=symbol]{siunitx} %[per-mode=reciprocal]
% Define additional units
\DeclareSIUnit\radians{rad}
\DeclareSIUnit\perunit{p.u.}
\DeclareSIUnit\var{VAr}

% Define units with brackets around it
\makeatletter
\newcommand{\sib}[1]{$\left[\si{#1}\right]$}
\makeatother

\makeatletter
\newcommand{\SIB}[2]{$\left[\SI{#1}{#2}\right]$}
\makeatother

% Define appearance of vectors and matrices. Default for \vec{} is an arrow
% \renewcommand{\vec}[1]{\mathbf{#1}} %bold
\renewcommand{\vec}[1]{#1} %nothing
% \let\oldvec\vec
% \renewcommand{\vec}[1]{\oldvec{\mathbf{#1}}} %bold with arrow
% \newcommand{\matr}[1]{\mathbf{#1}} %bold
\newcommand{\matr}[1]{#1} %nothing

% Define appearence of imaginary unit
%\newcommand{\iu}{\tilde{i}\mkern1mu} %non-slanted i, with dot. This seems to be the `official' guideline (ISO standard?)
% \newcommand{\iu}{\iota} %iota (has no dot on the i)
\newcommand{\iu}{\imath} %has no dot on the i but is curly

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Solving the power flow problem on integrated transmission-distribution networks: a review and numerical assessment}
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{~M.E.~Kootte,%~\IEEEmembership{Life~Fellow,~IEEE,}% <-this % stops a space
        ~J.E.~Romate,
        ~C.~Vuik%
% \thanks{M. Shell was with the Department
% of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
% GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
% \thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}%
}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%Field, gap, fill, terugkoppeling 
\noindent Power flow simulations form an essential tool for electricity network analysis but conventional models are designed to work on a separated transmission or distribution network only. The continuing growth of electricity consumption, demand side participation and renewable resources makes the electricity networks co-dependent. Integrated models incorporate the coupling of the networks and interaction that they have on each other, representing the power flow within this changing environment accurately. \\ 
Several numerical methods are available to solve the power flow problem on integrated networks. They can be categorized as a unified or as a splitting method and networks can be modeled as a homogeneous or hybrid network. In this paper, we review and assess these methods on the network models by running simulations on small test networks and comparing the outcome on accuracy, computational time, and convergence. The review shows that all the methods solve the power flow problem accurately. Methods running on hybrid networks have a slight advantage in computational time. Realistic network models, running on millions of buses and with large distribution networks, should give a better insight into the speed of the computations. 
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%nonlinear power flow, transmission-distribution, integrated networks, Newton-Raphson, convergence 
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle
\maketitle

%-------------
% Introduction
%-------------
\section{Introduction}
%Field, Gap, Focus 
Power flow computations are used to simulate the transport, generation, and consumption of power in electricity networks. The simulations are important for safe operation and planning of the electricity grid. An electricity grid consists of one transmission network, responsible for transport of power over large distances, and several distribution networks, responsible for the transport to end-consumers. Transmission networks transport high-voltage power to substations from where power is converted to a lower voltage level. The distribution network transports low-voltage power from substations to end-consumers. These transmission and distribution domains are separately analyzed where one domain is modeled in detail and the other is simplified in this model to one bus. Eg: in the analysis of the transmission system, the distribution network is simplified to one load in the transmission model. In the analysis of the distribution system,  the transmission network is simplified as a generator.  \newline\newline
The power system is changing: more renewable energy is entering the grid at distribution level, we see an increase of demand-side participation as a mechanism to balance frequency, and we see an increase of electricity consumption. The changing environment requires a more detailed analysis of the electricity network and a more complex model of the interaction between the networks. Integrated transmission-distribution network models study the interaction that these networks have on each other. But it is not straight-forward to integrate these separate domains. The networks have different characteristics what has resulted in different network models that cannot be easily integrated. The most important difference is that the transmission network is balanced and therefore modeled in single-phase while the distribution network is in general not balanced and modeled in three-phase. 
Furthermore, the different characteristics have lead to the development of different algorithms to solve the power flow problem. The power flow problem is a nonlinear problem which is solved using iterative methods. Well-known algorithms to solve transmission systems are: Gauss-Seidel, Newton-Raphson, Decoupled loadflow, and DC loadflow \cite{Schavemaker2008chap6}. These solvers have been adapted to distribution solvers, among which the most well-known are: Newton-Raphson current mismatch \cite{Garcia2000}\cite{Abdel-akher2005}, implicit $Z_{bus}$ \cite{Teng2002a}\cite{Chen1991a}, BIBC/BCBV \cite{Teng2008}\cite{Farag2011}, and Forward/Backward Sweep \cite{Moghaddas2009}\cite{Eminoglu2008} methods. Besides the technical complexity of integrating network models, we have the issue of confidentiality. System operators are not willing or not allowed to share network information with other system operators. \textit{One can reason that conventional models will at a certain moment not suffice anymore. \todo{De volgende zin moet er helemaal niet in of vriendelijker verwoord worden}Lagging legislation will then follow the necessary change towards integrating network models. 
}\newline\newline 
Integrated networks can be modeled as homogeneous and as hybrid networks. A homogeneous network model is a three-phase representation of both the transmission and distribution network. A hybrid network model consists of a single-phase transmission and a three-phase distribution network. Several numerical methods have been developed to solve integrated networks. They can be classified as a unified \cite{Taranto2008} or as a splitting method \cite{Sun2005}.  The unified methods solve the integrated system as a whole. The splitting method divides the networks into a master and a slave and defines an extra iterative scheme between them. At every iteration, the separated systems are solved on its own. The integrated system is solved once convergence on the boundary between the systems have been reached. This article reviews the proposed integration methods and analyzes them by running simulations on small test-cases. We pay special attention to their convergence behaviour: CPU-time and the number of iterations.

\section{Characterization of the power flow problem}
The steady-state power flow problem is the problem of determining the voltages $V$ in a network, given the specified power\footnote{We use the subscript $\iota$ as the imaginary unit.} $S=P+\iota Q$ and current $I$. $V$ and $I$ are related by Ohm's Law, $I = YV$, and $S$ and $V$ are related by $S=VI^*$. $Y$ represents the admittance of a power cable. Power is generated in three phases leading to three sinusoidal functions that describe phase $a$, $b$, and $c$ of the voltage and of the current. The voltage and current are both expressed in phasor notation: 
   \begin{align}
    V_p &=|V|_p\exp{(\iota\delta_{V_p}-\phi_{p})},\\
    I_p &=|I|_p\exp{(\iota\delta_{I_p}-\phi_{p})},\quad\ p\in\{a,b,c\},
    \end{align}
where $|\cdot|$ describes the phasor magnitude, $\delta_*$ the phase angle, and $\phi_*$ the phase shift. Phasors are often expressed as $V=\{|V|,\delta\}$. 
Current is never specified in an electricity system, therefore we substitute Ohm's Law into $S=VI^*$ and get a nonlinear equation for $S$, the three-phase nonlinear power flow equation, described as follows:  
\begin{align}
    S_p=V_p(\mathbf{Y}V)_p^*,\quad\ p\in\{a,b,c\}.
    \label{eq:SVYVp}
\end{align}
We represent an electricity network as a graph consisting of buses $i=1,..,N$ and branches (named after the two surrounding buses). These buses are either a load bus (PQ bus), a generator bus (PV bus), or a reference bus, depending on the information we know at that point. The loads in a network are modeled as PQ buses, loads consume power and at this bus, the active (P) and reactive power (Q) are specified.  Generators are modeled as PV buses, except for the first generator in a network, this bus is modeled as the slack bus. Generators supply power and at this bus, the active power (P) and voltage magnitude ($|V|$) are specified. At a slack bus, the voltage magnitude and angle are specified. This is summarized in table \ref{tab:networknodes}. 
\begin{table}[htbp]
\centering
\caption{Bustypes in a network and the information we know and not know at each bus $i$. }
\label{tab:networknodes}
\begin{tabular}{llllll}
\toprule
bus type  &&& \multicolumn{1}{c}{known} && \multicolumn{1}{c}{unknown} \\ \hline
$PQ$-bus  &&& $P_i,\ Q_i$               && $\delta_i,\ |V_i|$          \\
$PV$-bus  &&& $P_i,\ |V_i|$             && $Q_i,\ \delta_i$            \\
slack bus &&& $\delta_i,\ |V_i|$        && $P_i,\ Q_i$          \\      
\toprule
\end{tabular}
\end{table}
We scale all the engineering quantities in the power system to per-unit (pu) quantities by dividing them by their base value. In this way, we bring the voltage in a narrow range close to unity to eliminate erroneous values \cite{Schavemaker2008chap6}. \newline
Equation \eqref{eq:SVYVp} is a nonlinear equation so we solve it using iterative methods. At each node $i$, we solve the following equation for the unknown quantities:
\begin{align}
       S_i=V_i\sum_{k=1}^N{\mathbf{Y}}^*_{ik}{V}^*_k \label{eq:SVI}
\end{align}
The nodal admittance matrix $\mathbf{Y}_{ij}$ consists of admittance $y_{ij}$ of a line between node $i$ and $j$ and nodal shunt susceptance $b_c$. Line admittance consists of a real and imaginary part: $y_{ij}= 1/z_{ij}= 1 / (r_{ij}+\iota x_{ij})$, $z$ being impedance, $r$ being resistance and $x$ being reactance.  The nodal admittance matrix relating node $i$ and $j$ looks as the following: 
\begin{align}\mathbf{Y}_{ij}= 
\begin{bmatrix}
(y_{ij}+\iota \frac{b_c}{2})\frac{1}{\tau^2} & -y_{ij}\frac{1}{\tau\exp{(-\iota\theta_s)}}\nonumber \\
-y_{ij}\frac{1}{\tau\exp{(\iota\theta_s)}}& y_{ij}+\iota \frac{b_c}{2}   
\end{bmatrix} = \begin{bmatrix} Y_{11} & Y_{12} \\ Y_{21} & Y_{22}
\end{bmatrix}_{ij}
\end{align}
The parameters $\theta_s$ and $\tau$ are the transformer's phase-shift angle and tap ratio, when there is no transformer between two nodes but just a normal power cable, $\theta_s=0$ and $\tau=1$.

>>>>

\subsubsection{Accuracy} 
To check the accuracy of the methods, we compare the voltage output of the integrated methods with the output of the separated methods. We solve the transmission network separately using NR-power and we solve the distribution network separately using NR-TCIM. The distribution domain in the integrated network has no reference bus anymore. In order to compare the voltage profile of this part, we scale the values accordingly. 
We know that the voltage is given by: \begin{equation}
V_p=|V|_p\exp{(\iota\delta_{V_p}-\phi)},\quad p\in\{a,b,c\},
\end{equation}
where $|V|$ and $\delta$ are set to $|V|=1.0$ pu and $\delta=0$ at the reference bus. We can compare the voltage of the distribution domain to the voltage of the separated domain after scaling. We divide the voltage magnitude of the distribution domain, buses $i=1,...,N_D$, by the voltage magnitude of the distribution connection bus: 
\begin{equation}
|V|_{i,{D_{new}}} = \frac{|V|_{i,{D_{uns}}}}{|V|_{ref,{D_{uns}}}},\quad i=1,..,N_D.
\label{eq:ViDnew}
\end{equation}
The subscript $uns$ represents the unscaled values and $new$ represents the scaled values. \newline 
The voltage angle represents the phase-shift in relation to the phase of the reference bus. In order to compare the integrated output of the voltage angle with the separated output, we subtract the phase-angle of the former reference bus from all the buses $i=1,..,N_D$: 
\begin{equation}
\delta_{i,{D_{new}}} ={\delta_{i,{D_{uns}}}}-{\delta_{ref,{D_{uns}}}},\quad i=1,..,D_N.
\label{eq:dViDnew}
\end{equation}
Note that we need to apply \eqref{eq:ViDnew} and \eqref{eq:dViDnew} to all the three phases separately. \newline\newline
We compare the voltage magnitudes and angles of the integrated network, including the scaled distribution domain, with the magnitudes angles of the separate domains. We take the infinity norm of the relative difference between the voltages: 
\begin{equation}
\mbox{rel. error of } |V|_p = \left|\frac{|V|^*_p - |V|^o_p}{|V|^o_p}\right|_\infty,\quad p\in\{a,b,c\},
\label{eq:norm|V|}
\end{equation}
where the asterisk $*$ marks the outcome of separated networks, and the superscript $o$ the output of separated networks (o=original). We do this in a similar manner for the voltage angle: 
\begin{equation}
\mbox{rel. error of } \delta_p = \left|\frac{\delta^*_p - \delta^o_p}{\delta^o_p}\right|_\infty,\quad p\in\{a,b,c\},
\label{eq:normdelta}
\end{equation}
We have given the relative errors of the voltage magnitude and angle of the three phases for all the test cases in table \ref{tab:relerror}.
